# Tidyplots from time to time

## Font size in mm or pt

> ... **ggplot2 specifies the size in millimeters (mm)**, rather than the usual points (pts). The reason for this choice is that it makes it the units for font sizes consistent with how other sizes are specified in ggplots. ([**There are 72.27 pts in a inch, so to convert from pts to mm, just multiply by 25.4 / 72.27**]{.mark}).

From [ggplot2: Elegant Graphics for Data Analysis (3e)](https://ggplot2-book.org/annotations.html#sec-text-labels)

Since 1 inch = 2.54 cm = 25.4 mm, and there are 72.27 pts in 25.4 mm,

so there are 2.845276 pts in a mm,

```{r}
72.27 / 25.4
```

or one pt equals to 0.3514598 mm.

```{r}
25.4 / 72.27
```

If it is needed to convert from mm to pts, just: mm * (72.27 / 25.4); or from pts to mm, just: pts * (25.4 / 72.27).

There is a numeric (i.e. `.pt`) in ggplot2. The numeric is equal to `72.27 / 25.4`.

```{r}
mode(ggplot2::.pt)

ggplot2::.pt == 72.27 / 25.4

round(1 / ggplot2::.pt, digits = 7) == round(25.4 / 72.27, digits = 7)

# For example
12 / ggplot2::.pt
```

In other words, if it is needed to convert from mm to pts, just: mm * `ggplot2::.pt`; or from pts to mm, just: pts * (1 / `ggplot2::.pt`) or pts / `ggplot2::.pt`.

```{r}
#| message: false
library(tidyplots)

df <- tibble::tibble(x = 1, y = 1)

df

df |>
  tidyplot(x = x, y = y) |> 
  add_annotation_text(
    text = "12 pts or 4.22 mm",
    x = 0.75,
    y = 1,
    color = "#bb5566",
    fontsize = 12) |> # default in pt
  add(ggplot2::annotate(
    "text", 
    x = 1.25, 
    y = 1, 
    label = "12 pts or 4.22 mm",
    color = "#004488",
    size = 12 / ggplot2::.pt)) |> # convert 12 pts to the default mm (4.217518)
  add_title(title = "12 pts is approximately 4.22 mm\nin fontsize") |> 
  adjust_x_axis(limits = c(0.5, 1.5)) |> 
  adjust_size(width = 100)
```

## Could `add_heatmap()` add dendrogram for rows or columns?

The answer is `NO`, since `add_headmap()` is a wrap of `ggplot2::geom_raster`.

```{r}
library(tidyplots)

add_heatmap

ggplot2::faithfuld |> 
  dplyr::slice_head(n = 10)

ggplot2::faithfuld |> 
  tidyplot(x = waiting, y = eruptions, color = density) |> 
  add_heatmap()
```

So sometimes we need to pre-order the rows and/or columns before using `add_headmap()`.