# Tidyplots from time to time

## Font size in mm or pt

> ... **ggplot2 specifies the size in millimeters (mm)**, rather than the usual points (pts). The reason for this choice is that it makes it the units for font sizes consistent with how other sizes are specified in ggplots. ([**There are 72.27 pts in a inch, so to convert from pts to mm, just multiply by 25.4 / 72.27**]{.mark}).

From [ggplot2: Elegant Graphics for Data Analysis (3e)](https://ggplot2-book.org/annotations.html#sec-text-labels)

Since 1 inch = 2.54 cm = 25.4 mm, and there are 72.27 pts in 25.4 mm,

so there are 2.845276 pts in a mm,

```{r}
72.27 / 25.4
```

or one pt equals to 0.3514598 mm.

```{r}
25.4 / 72.27
```

If it is needed to convert from mm to pts, just: mm * (72.27 / 25.4); or from pts to mm, just: pts * (25.4 / 72.27).

There is a numeric (i.e. `.pt`) in ggplot2. The numeric is equal to `72.27 / 25.4`.

```{r}
mode(ggplot2::.pt)

ggplot2::.pt == 72.27 / 25.4

round(1 / ggplot2::.pt, digits = 7) == round(25.4 / 72.27, digits = 7)

# For example
12 / ggplot2::.pt
```

In other words, if it is needed to convert from mm to pts, just: mm * `ggplot2::.pt`; or from pts to mm, just: pts * (1 / `ggplot2::.pt`) or pts / `ggplot2::.pt`.

```{r}
#| message: false
library(tidyplots)

df <- tibble::tibble(x = 1, y = 1)

df

df |>
  tidyplot(x = x, y = y) |> 
  add_annotation_text(
    text = "12 pts or 4.22 mm",
    x = 0.75,
    y = 1,
    color = "#bb5566",
    fontsize = 12) |> # default in pt
  add(ggplot2::annotate(
    "text", 
    x = 1.25, 
    y = 1, 
    label = "12 pts or 4.22 mm",
    color = "#004488",
    size = 12 / ggplot2::.pt)) |> # convert 12 pts to the default mm (4.217518)
  add_title(title = "12 pts is approximately 4.22 mm\nin fontsize") |> 
  adjust_x_axis(limits = c(0.5, 1.5)) |> 
  adjust_size(width = 100)
```

## Could `add_heatmap()` add dendrogram for rows or columns?

The answer is `NO`, since `add_headmap()` is a wrap of `ggplot2::geom_raster`.

```{r}
library(tidyplots)

add_heatmap

ggplot2::faithfuld |> 
  dplyr::slice_head(n = 10)

ggplot2::faithfuld |> 
  tidyplot(x = waiting, y = eruptions, color = density) |> 
  add_heatmap()
```

So sometimes we need to pre-order the rows and/or columns before using `add_headmap()`.

## How long have I lived in different places?

When I went for a walk with my wife this afternoon, our chat turned to the question: how long have I spent in different places?

So let me calculate it here.

```{r}
#| message: false
library(tidyplots)
library(lubridate)

current_ym <- format(Sys.Date(), "%Y-%m")

df <- tibble::tibble(
  places = factor(
    x = c("Chengcheng county", "Yangling town", "Shanghai"),
    levels = c("Chengcheng county", "Yangling town", "Shanghai"),
    ordered = TRUE
  ),
  days = as.numeric(c( # the time is not so precise (i.e. not to days)
    ym("2001-08") - ym("1983-02"),
    ym("2007-07") - ym("2001-09"),
    ym(current_ym) - ym("2007-08")
  ))
)

df

df |> 
  tidyplot(x = places, y = days, color = places) |> 
  add_sum_bar() |> 
  add_sum_value() |> 
  add_caption(caption = paste("as of", current_ym)) |> 
  adjust_x_axis(rotate_labels = TRUE)
```

I fist lived in Chengcheng county (Shaanxi, China), followed by Yangling town (Shaanxi, China). I am currently living in Shanghai.

So I have lived in Shanghai the longest.

## How to understand the logic of `downsample` parameter of `adjust_colors` function?

This is a source code snippet of `tidyplots::adjust_colors`:

```{r}
#| eval: false
        if (n_ratio >= 1) { # n_ratio refers to n_provided/n_requested
            if (n_ratio > 1) {
                new_colors <- downsample_vector(new_colors, n_requested, 
                  downsample = downsample)
            }
            suppressMessages(out <- out + ggplot2::scale_color_manual(values = new_colors, 
                drop = TRUE, labels = labels, ...))
            suppressMessages(out <- out + ggplot2::scale_fill_manual(values = apply_saturation(new_colors, 
                saturation = effective_saturation), drop = TRUE, 
                labels = labels, ...))
        }
```

So we need to check the function of `tidyplots:::downsample_vector`:

```{r}
downsample_vector <- function (x, n, downsample = c("evenly", "first", "last", "middle")) 
{
    if (length(x) <= n) {
        return(x)
    }
    downsample <- match.arg(downsample)
    if (downsample == "evenly") {
        by <- (length(x)/(n - 1)) - (1/(n - 1))
        i <- floor(cumsum(c(1, rep(by, n - 1))))
        x[i]
    }
    else if (downsample == "first") {
        x[1:n]
    }
    else if (downsample == "last") {
        x[(length(x) - n + 1):length(x)]
    }
    else {
        start_index <- ceiling((length(x) - n)/2) + 1
        end_index <- start_index + n - 1
        x[start_index:end_index]
    }
}
```

### How about n = 1 (i.e. n_provided > n_requested)?

```{r}
# New color scheme with length 3
new_colors <- c("red", "blue", "black")
new_colors
```

```{r}
downsample_vector(new_colors, 1, downsample = "evenly")
downsample_vector(new_colors, 1, downsample = "first")
downsample_vector(new_colors, 1, downsample = "last")
downsample_vector(new_colors, 1, downsample = "middle")
```

### How about n = 2 (i.e. n_provided > n_requested)?

```{r}
downsample_vector(new_colors, 2, downsample = "evenly")
downsample_vector(new_colors, 2, downsample = "first")
downsample_vector(new_colors, 2, downsample = "last")
downsample_vector(new_colors, 2, downsample = "middle")
```

### How about n = 3 (i.e. n_provided = n_requested)?

```{r}
downsample_vector(new_colors, 3, downsample = "evenly")
downsample_vector(new_colors, 3, downsample = "first")
downsample_vector(new_colors, 3, downsample = "last")
downsample_vector(new_colors, 3, downsample = "middle")
```

### How about n = 4 (i.e. n_provided < n_requested)?

```{r}
downsample_vector(new_colors, 4, downsample = "evenly")
downsample_vector(new_colors, 4, downsample = "first")
downsample_vector(new_colors, 4, downsample = "last")
downsample_vector(new_colors, 4, downsample = "middle")
```

## `dnorm`, `pnorm`, `qnorm`, and `rnorm`?

My math and statistics are, to be honest, not good, and neither is my R. Here I want to understand the four functions outlined in the section title.

`dnorm`: density for the normal distribution; `pnorm`: distribution function for the normal distribution; `qnorm`: quantile function for the normal distribution; `rnorm`: random generation for the normal distribution.

```{r}
library(tidyplots)

# dnorm
plot_dnorm <- tibble::tibble(
  d_norm = dnorm(x = seq(-5, 5, by = 0.05)), # x: quantiles
  x = seq(-5, 5, by = 0.05)
)
  
p1 <- plot_dnorm |>
  tidyplot(x = x, y = d_norm) |> 
  add_data_points(alpha = 0.2) |> 
  add_title(title = "'dnorm': density")

# pnorm
plot_pnorm <- tibble::tibble(
  p_norm = pnorm(q = seq(-5, 5, by = 0.05)), # q: quantiles
  x = seq(-5, 5, by = 0.05)
)

p2 <- plot_pnorm |> 
  tidyplot(x = x, y = p_norm) |> 
  add_data_points(alpha = 0.2) |> 
  add_title(title = "'pnorm': distribution")

# qnorm
plot_qnorm <- tibble::tibble(
  q_norm = qnorm(p = seq(0, 1, by = 0.005)), # p: probability
  x = seq(0, 1, by = 0.005)
)

p3 <- plot_qnorm |> 
  tidyplot(x = x, y = q_norm) |> 
  add_data_points(alpha = 0.2) |> 
  add_title(title = "'qnorm': quantile")

# rnorm
plot_rnorm <- tibble::tibble(
  r_norm = rnorm(n = 1000) # n: number of observations
)

p4 <- plot_rnorm |> 
  tidyplot(x = r_norm) |> 
  add_histogram(bins = 100) |> 
  add_title(title = "'rnorm': random")

patchwork::wrap_plots(p1, p2, p3, p4, ncol = 2)
```

Based on the plots, I (and might you) can now understand them intuitively.

## Use `Reduce` for functional programming

This is the [parallel version](https://ggplot2-book.org/programming.html#exercises-3) but applying to `tidyplots`.

```{r}
library(tidyplots)

add_points_curve <- function(x, y, ...) {
  Reduce(
    f = function(init_plot, next_component) {next_component(init_plot)},
    x = list(...),
    init = ggplot2::mpg |> 
      tidyplot(x = {{x}}, y = {{y}}) |> 
      add_title(title = "p0"),
    accumulate = TRUE
  )
}

add_points_curve(
  x = displ,
  y = hwy,
  function(plot) {
    add_data_points_beeswarm(plot, alpha = 0.4) |> 
    adjust_title(title = "p1")
  },
  function(plot) {
    add_curve_fit(plot, fill = "#bb5566") |> 
    adjust_title(title = "p2")
  }
) |> 
  patchwork::wrap_plots(ncol = 3)
```